---
title: "Preliminary occurance check and Maxent model for for V. velutina & V. mandarina"
author: "Daniel Flø"
date: "`r Sys.Date()`"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

<br>

# Conclusion: there is enough data to predict species potential distributions.

<br>


```{r setup, include=FALSE}

require(knitr)
knitr::opts_knit$set(root.dir = 'C:/Users/DAFL/Dropbox/vkm/Vespa velutina/Vespa velutina/')

library(spocc)
library(rgbif)
library(tidyverse)
library(taxize)
library(dplyr)
library(scrubr)
library(foreach)
library(doParallel)
library(raster)
library(dismo)

```

<br>

######  1 gnr_resolve Resolve names using Global Names Resolver
######  2 get_gbifid_ include synonymes 
######  3 gbif
######  4 occ multiple databases 
######  5 plot 

<br>

```{r}

splist  <- c("Vespa velutina",  "Vespa mandarina")

```

<br>

```{r pressure, echo=FALSE, warning = FALSE}

# Resolve names using Global Names Resolv
species<-taxize::gnr_resolve(splist, best_match_only=T)

# make new vektor of species to check gbif backbone occurance 
splist<-species$matched_name

# fill in your gbif.org credentials 
user <- "danief"  
pwd <- "ZXasqw12" 
email <- "floedaniel@gmail.com" 

splist<- splist %>% 
  taxize::get_gbifid_(method="backbone") %>% # match names to the GBIF backbone to get taxonkeys
  purrr::imap(~ .x %>% mutate(original_sciname = .y)) %>% # add original name back into data.frame
  bind_rows() 

```

<br>

```{r splist}

splist

```

<br>

```{r spocc, echo=FALSE, warning = FALSE}


library(spocc)
## Loading required package: ggplot2
out <- occ(query = splist$scientificname, from =c("gbif", "bison", "obis", "ala", "inat", "idigbio"), limit = 9000)

df_out <- occ2df(out)

as_tibble(df_out)

unique(df_out$prov)

df_out <- dplyr::select(df_out, name, longitude, latitude)

names(df_out)<-c("Species","Longitude", "Latitude")

df_out$Longitude <- as.numeric(df_out$Longitude)
df_out$Latitude <- as.numeric(df_out$Latitude)

```

<br>

```{r bind, echo=FALSE}

df <- df_out

```

<br>

```{r basemap, echo=FALSE}

wrld <- map_data("world") %>% filter(!region=="Antarctica")

```

<br>

##  Plot all data

<br>

```{r plot_one, fig.dim = c(10, 10), warning = FALSE}

ggplot() +
  geom_map(data=wrld, map=wrld, aes(map_id=region , x=long, y=lat), fill="black", col="white") +
  geom_point(data=df, aes(x=Longitude, y=Latitude, col=Species)) +
  coord_quickmap() +
  theme_void()

```

<br>
<br>


```{r clean, echo=FALSE, warning = FALSE}

df <-df_out %>%
  coord_impossible() %>%
  coord_incomplete() %>% 
  coord_unlikely()

df <- df %>% coord_unlikely() 

df$Species <- tolower(df$Species)
df <- dplyr::filter(df, !grepl("BOLD", Species))
df$Species <- word(df$Species, start=1, end=2)

```

<br>

##  Plot clean data (i.e no dups or impossible)

<br>

```{r plot_clean, fig.dim = c(10, 10), warning = FALSE}

ggplot() +
  geom_map(data=wrld, map=wrld, aes(map_id=region , x=long, y=lat), fill="black", col="white") +
  geom_point(data=df, aes(x=Longitude, y=Latitude, col=Species)) +
  coord_quickmap() +
  theme_void() 

```

<br>

##  Check how many in Asia

<br>

```{r in_asia}

df %>%  filter(Longitude > 60) %>% count()

```

<br>

##  Plot only occurence in Asia 

<br>

```{r plot_asia, fig.dim = c(10, 10), warning = FALSE}

ggplot() +
  geom_map(data=wrld, map=wrld, aes(map_id=region , x=long, y=lat), fill="black", col="white") +
  geom_point(data=filter(df, Longitude > 60), aes(x=Longitude, y=Latitude, col=Species)) +
  coord_quickmap() +
  theme_void() 


```

<br>

##  Download predictors

<br>

```{r predictors,  warning = FALSE}

# Bring in the env layers from the dismo package
# predictors <- raster::getData("worldclim", var="bio", res=10)

library(rnaturalearth)
library(rnaturalearthdata)

eurasia <- ne_countries(continent = c('europe', "asia"), scale = "medium", returnclass = "sp" )
extent(bbox(eurasia))
extent <- c(-25, 180, -21.36904, 81.8542)

library(raster)

dir.create("C:/Users/DAFL/Dropbox/vkm/Vespa velutina/Vespa velutina/Study Region/", recursive=TRUE, showWarnings=FALSE)

library(doParallel)  #Foreach Parallel Adaptor 
library(foreach)     #Provides foreach looping construct

# Define how many cores you want to use
UseCores <- detectCores() -1
# Register CoreCluster
cl       <- makeCluster(UseCores)
registerDoParallel(cl)

stack_list <- list.files(path = "C:/Users/DAFL/Desktop/worldclim", full.names = TRUE, pattern = "\\.tif$")

# for each BIOCLIM raster
#Use foreach loop and %dopar% command
foreach(i=1:length(stack_list), .packages = c("raster")) %dopar% {
  

  files <- list.files(path = "C:/Users/DAFL/Desktop/worldclim", full.names = TRUE, pattern = "\\.tif$")
  
  rast <- raster(files[i])
  
  filename <- (paste(basename(files[i]), sep = ""))
  
  # crop the raster and update tt, 
  rast <- setMinMax(rast)
  projection(rast) <- '+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0'
  
  # if (i %in% c(1:2, 5:11)) rast <- rast / 10
  
  # Crop the raster
  rast <- crop(rast, extent, snap = "out")
  
  
  rast<- mask(rast, eurasia)
  
  # save raster
  writeRaster(rast,
              paste0('./Study Region/', filename, i), 
              format='GTiff',
              datatype='INT2S',
              overwrite = TRUE)
  
  
}

#end cluster
stopCluster(cl)

# visually check
predictors <- stack(
  list.files(
    'C:/Users/DAFL/Dropbox/vkm/Vespa velutina/Vespa velutina/Study Region/',
    full.names=TRUE,
    pattern='.tif'
  )
)


```

<br>

##  Reshape and filter

<br>

```{r select,  warning = FALSE}

# must have this structure 
df <- df %>% dplyr::select(x=Longitude, y=Latitude, Species=Species)

df <- df %>%  group_by(Species) %>%  filter(n() > 50)

df<-as.data.frame(df)

```


<br>

##  Run Maxnt model using with multicore. Untuned model with default settings! All 20 variables

<br>

```{r dopar,  warning = FALSE}

setwd("C:/Users/DAFL/Dropbox/vkm/Vespa velutina/Vespa velutina/")

# create list of unique species to loop over 
sp_list <- unique(df$Species)

numCores <- detectCores()
numCores
 # use multicore, set all
registerDoParallel(numCores) 

max <- foreach (i=seq_along(sp_list), .packages = c("dismo", "rJava")) %dopar% {
  
  data=subset(df, df$Species==sp_list[i])
  
  occ <- data[, -3]
  
  # witholding a 25% sample for testing
  fold <- kfold(occ, k = 5)
  occtest <- occ[fold == 1, ]
  occtrain <- occ[fold != 1, ]
  
  me <- maxent(predictors, occtrain)
  
  # project into geographic space
  r <- predict(me, predictors, args=c("outputformat=cloglog", "betamultiplier=1.5") )
  
}

doParallel::stopImplicitCluster()

```

<br>

##  Stack raster

<br>

```{r stack,  warning = FALSE}

# stack all the outputs and name them
stack.out <- stack(max)
names(stack.out) <- sp_list
#plot(stack.out)

```

<br>

##  Plot MaxEnt result 
## cloglog output =  a zero to one probability of presence for the species 

<br>

```{r plot_loop, fig.dim = c(10, 10), warning = FALSE}

for (i in 1:nlayers(stack.out)) {
  
  p=subset(df, df$Species==sp_list[i])
  
  plot(stack.out[[i]], main=sp_list[i], col=rev(terrain.colors(10, alpha = 1)), breaks = seq(0, 1, by = 0.1)) 
  
  #points(p[,1:2], type = "p", pch=16, size=1)
  
}

```


```{r plot_gg, fig.dim = c(10, 10), warning = FALSE}

library(viridis)

mandarina_pts <- rasterToPoints(stack.out$vespa.mandarinia, spatial = TRUE)
mandarina_df  <- data.frame(mandarina_pts)

ggplot() +
  geom_raster(data = mandarina_df , aes(x = x, y = y, fill = vespa.mandarinia  )) + 
  scale_fill_gradientn(colours=plasma(7),  breaks=c(0, 0.5, 1), labels=c(0 ,0.5, 1), limits=c(0, 1)) +
  coord_cartesian(xlim = c(0, 30), ylim = c(58, 71))  + 
  labs(title = "vespa.mandarinia",
       subtitle = "Current climate - Bioclim predictors")

ggplot() +
  geom_raster(data = mandarina_df , aes(x = x, y = y, fill = vespa.mandarinia  )) + 
  scale_fill_gradientn(colours=plasma(7),  breaks=c(0, 0.5, 1), labels=c(0 ,0.5, 1), limits=c(0, 1)) +
  coord_cartesian(xlim = c(-20, 50), ylim = c(30, 71))  + 
  labs(title = "vespa.mandarinia",
       subtitle = "Current climate - Bioclim predictors")


ggplot() +
  geom_raster(data = mandarina_df , aes(x = x, y = y, fill = vespa.mandarinia  )) + 
  scale_fill_gradientn(colours=magma(3),  breaks=c(0, 0.5, 1), labels=c(0 ,0.5, 1), limits=c(0, 1)) + 
  geom_point(data = filter(df, Species=="vespa mandarinia"), aes(x=x, y=y), col="red") +
  coord_cartesian(xlim = c(-20, 50), ylim = c(30, 71))  + 
  labs(title = "vespa.mandarinia",
       subtitle = "Current climate - Bioclim predictors")


#####

vespa.velutina_pts <- rasterToPoints(stack.out$vespa.velutina, spatial = TRUE)
vespa.velutina_df  <- data.frame(vespa.velutina_pts)

ggplot() +
  geom_raster(data = vespa.velutina_df , aes(x = x, y = y, fill = vespa.velutina  )) + 
  scale_fill_gradientn(colours=plasma(7),  breaks=c(0, 0.5, 1), labels=c(0 ,0.5, 1), limits=c(0, 1)) +
  coord_cartesian(xlim = c(0, 30), ylim = c(58, 71))  + 
  labs(title = "vespa.velutina",
       subtitle = "Current climate - Bioclim predictors")

ggplot() +
  geom_raster(data = vespa.velutina_df , aes(x = x, y = y, fill = vespa.velutina   )) + 
  scale_fill_gradientn(colours=plasma(7),  breaks=c(0, 0.5, 1), labels=c(0 ,0.5, 1), limits=c(0, 1)) +
  coord_cartesian(xlim = c(-20, 50), ylim = c(30, 71))  + 
  labs(title = "vespa.velutina",
       subtitle = "Current climate - Bioclim predictors")



ggplot() +
  geom_raster(data = vespa.velutina_df , aes(x = x, y = y, fill = vespa.velutina   )) + 
  scale_fill_gradientn(colours=magma(3),  breaks=c(0, 0.5, 1), labels=c(0 ,0.5, 1), limits=c(0, 1)) +
    geom_point(data = filter(df, Species=="vespa velutina"), aes(x=x, y=y), col="red") +
  coord_cartesian(xlim = c(-20, 50), ylim = c(30, 71))  + 
  labs(title = "vespa.velutina",
       subtitle = "Current climate - Bioclim predictors & presence points")



```
